NATIONALISTIC BIAS IN SPORT PERFORMANCE EVALUATIONS

### 2020-08-02 accepted-ms/second_response_to_reviewers.Rmd

---
title: "Response to EJSS reviewers' comments on manuscript #2019-0005-R1"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-memo.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
csl: /home/torkildl/Dropbox/Literature/csl-files/american-sociological-review.csl
anonymous: yes
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

We are grateful for being granted the opportunity to revise and resubmit our manuscript. We have considered all the comments by both the reviewers and the Editor.

In what follows, we respond to each of the comments and explain how we accordingly have adapted our work.

Thank you for reading and commenting on our work!

# Responses to comments made by reviewer \#3

> I congratulate the authors for their revision and recommend to accept the paper with minor changes.

> First of all I'd like to thank the authors for seriously considering the recommendations, which were certainly not easily to address from their initial perspective. Especially by incorporating more theoretical literature and applying it to their case, authors have made the paper much more accessible for readers of the EJSS. Of course, the theoretical framework still is not highly elaborate, but now there are some interesting and sensible suggestions how the topic and findings can be theorized. Thus the discussion has gained more substance, too. And the integration of more recent empirical data makes the analysis more convincing.

>Apart from adressing the few corrections I have made in the pdf file, authors should check the new sections (p.5ff., 18ff.) for grammatical errors, or maybe let a native speaker revise those parts.

We have re-read and corrected the relevant sections for typos.

# Responses to comments made by reviewer \#4

>I like the revision. The passages to which the critical comments of Reviewer 2 referred have, in my opinion, been appropriately revised. I recommend to accept the paper with some minor changes:

>Firstly, the authors use the term "nationalist bias" in the abstract and "nationalistic bias" in the text. I would recommend that the authors choose one of the two, preferably "nationalistic bias", as this term is (to my knowledge) more often used to describe actions.

We now consistently refer to nationalistic bias throughout the manuscript.

>Secondly, I would recommend to avoid using the terms "cultural" and "national" in the same context as on page 3. National and cultural are two different things and should not be confused conceptually. Since the text, as I understand it, is about a nationalistic bias, I would delete the terms "cultural" and "culturally" on page 3.

We have rewritten the text and now avoid confusing nationalistic and cultural biases.

>Thirdly, there are several more types in the text. I would recommend that the revision be checked again grammatically and orthographically by a native speaker.

We have re-read and corrected typos found throughout the manuscript.

\medskip

\noindent
Thank you again to reviewers and Editors for comments on our manuscript!

--The Authors

\bigskip

END.

### 2020-08-02 accepted-ms/manuscript.Rmd
Accepted manuscript update

---
title: "NATIONALISTIC BIAS IN SPORT PERFORMANCE EVALUATIONS"
subtitle: "An example from the ski jumping World Cup"
runhead: Judging biases in skijumping
author:
- affiliation: Department of Sociology and Human Geography, University of Oslo.
  name: Torkild Hovde Lyngstad
- affiliation: European University Institute
  name: Juho Härkönen
- affiliation: Department of Medical Statistics, University of Oslo
  name: Leiv Tore Salte Rønneberg
date: "14 september 2017"
keywords: nationalistic bias, skijumping, judges, fixed effects
abstract: "Illegitimate biases of non-independent judges represent a threat to the legitimacy of international sport competitions. Judges’ nationalities represent one source of bias, as they consciously or subconsciously may prefer athletes from their own nation. Such biases may affect the outcomes of competitions. The literature offers no complete consensus on the magnitude, origin and stability of such biases. In this paper, we shed light on these problems, using international ski jumping as an example. We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons and estimate a series of fixed-effects models to test hypotheses on nationalistic biases. Our results reaffirm suspicions of nationalistic bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes. The biases vary between nations, but do not change markedly over the study period."
thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.  At publication time, all code and material used to produce this paper (with the exception of the proprietary data) will be available. The authors are also grateful to Matti Nykänen for being an enduring source of inspiration in many aspects of our lives."
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-ms.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
anonymous: no
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

@@

# Our results in light of the extant literature

>>>>

---
title: "NATIONALISTIC BIAS IN SPORT PERFORMANCE EVALUATIONS"
subtitle: "An example from the ski jumping World Cup"
runhead: Judging biases in skijumping
author:
- affiliation: Department of Sociology and Human Geography, University of Oslo.
  name: Torkild Hovde Lyngstad
- affiliation: European University Institute
  name: Juho Härkönen
- affiliation: Department of Medical Statistics, University of Oslo
  name: Leiv Tore Salte Rønneberg
date: "14 september 2017"
keywords: nationalistic bias, skijumping, judges, fixed effects
abstract: "Illegitimate biases of non-independent judges represent a threat to the legitimacy of international sport competitions. Judges’ nationalities represent one source of bias, as they consciously or subconsciously may prefer athletes from their own nation. Such biases may affect the outcomes of competitions. The literature offers no complete consensus on the magnitude, origin and stability of such biases. In this paper, we shed light on these problems, using international ski jumping as an example. We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons and estimate a series of fixed-effects models to test hypotheses on nationalistic biases. Our results reaffirm suspicions of nationalistic bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes. The biases vary between nations, but do not change markedly over the study period."
thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.  At publication time, all code and material used to produce this paper (with the exception of the proprietary data) will be available. The authors are also grateful to Matti Nykänen for being an enduring source of inspiration in many aspects of our lives."
output:
  word_document:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-ms.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
anonymous: no
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

@@

## Our results in light of the extant literature
